Reliability-Aware Hierarchical Learning for Chagas Detection from Electrocardiogram under Expert Label Scarcity

### 2025-08-28 poster.tex
Complete poster.tex for 2025 Chagas ECG screening challenge
Co-authored-by: wenh06 <8778305+wenh06@users.noreply.github.com>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% NIWeek 2014 Poster by T. Reveyrand
% www.microwave.fr
% http://www.microwave.fr/LaTeX.html
% ---------------------------------------
%
% Original template created by:
% Brian Amberg (baposter@brian-amberg.de)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\usepackage{textcomp}

\usepackage{hyperref}

@@

 \definecolor{bordercol}{RGB}{40,40,40} % Border color of content boxes
 \definecolor{headercol1}{RGB}{186,215,230} % Background color for the header in the content boxes (left side)
 \definecolor{headercol2}{RGB}{120,120,120} % Background color for the header in the content boxes (right side)
 \definecolor{headerfontcol}{RGB}{0,0,0} % Text color for the header text in the content boxes
 \definecolor{boxcolor}{RGB}{210,235,250} % Background color for the content in the content boxes

\definecolor{jdhblue}{RGB}{2,93,186}

\definecolor{zkblue}{RGB}{88,135,175}
\definecolor{zkbackground}{RGB}{230,232,234}

% \usetikzlibrary{shapes, arrows, external, decorations.pathmorphing, backgrounds, positioning, fit, petri, calc, hobby, cd}

@@

% \background{ % Set the background to an image (background.pdf)
% \begin{tikzpicture}[remember picture,overlay]
% \draw (current page.north west)+(-2em,2em) node[anchor=north west]
% {\includegraphics[height=1.1\textheight]{images/baposter_background.pdf}};
% \end{tikzpicture}
% }

\begin{poster}{
grid=false,
borderColor=bordercol, % Border color of content boxes
headerColorOne=headercol1, % Background color for the header in the content boxes (left side)
headerColorTwo=headercol2, % Background color for the header in the content boxes (right side)
headerFontColor=headerfontcol, % Text color for the header text in the content boxes
% boxColorOne=boxcolor, % Background color for the content in the content boxes
boxColorOne=zkbackground,
headershape=roundedright, % Specify the rounded corner in the content box headers
headerfont=\Large\sf\bf, % Font modifiers for the text in the content box headers
textborder=rectangle,
% background=user,
background=none,
headerborder=open, % Change to closed for a line under the content box headers
boxshade=plain
}
%
%----------------------------------------------------------------------------------------
%	TITLE AND AUTHOR NAME
%----------------------------------------------------------------------------------------
%
{\includegraphics[scale=0.401]{images/logo_cau_science.png}} % University/lab logo
{
{\bf \fontsize{19pt}{19pt} \selectfont To write}
} % Poster title
{\vspace{0.3em} \smaller Hao WEN$^1$, Jingsu KANG$^2$  \\  % Author names

$^1${\it College of Science, China Agricultural University}\qquad
$^2${\it Tianjin Medical University} \\
\vspace{0.2cm}
{\Large \bf{The George B. Moody PhysioNet Challenge 2025}, ~~~\bf{Team Revenger}}
}
{\includegraphics[scale=0.101]{logo_tmu.jpeg}}

\newif\ifcoloredtext
\coloredtexttrue
\newif\ifboxednn
\boxednntrue

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\headerbox{Introduction}{name=introduction, column=0, row=0, span=3}{
% finished

\begin{itemize}
\item to write
\vspace{-0.2cm}
\item to write
\vspace{-0.2cm}
\item to write
\end{itemize}
}

%----------------------------------------------------------------------------------------
%	Framework
%----------------------------------------------------------------------------------------

\headerbox{The Multi-Stage Framework}{name=framework, column=0, span=3, row=1, below=introduction}{
% finished

\newif\ifFrameworkFigBox
\FrameworkFigBoxfalse

\begin{center}
\input{tikz_plots/multi-stage-framework}
\end{center}

@@

%----------------------------------------------------------------------------------------
%	Preprocessing
%----------------------------------------------------------------------------------------

\headerbox{Training Data Preparation}{name=data, column=0, row=1, span=2, below=framework}{
% finished

\begin{itemize}
\item ECG-Image-Kit to generate ECG images from the PTB-XL dataset for our method development. Additionally, the PTB-XL+ dataset was used to generate classification labels.
\vspace{-0.2cm}
\item Image generation setup included augmentations such as random Gaussian noise, random color temperature adjustments, and distortions including handwritten notes, wrinkles, and creases.
\vspace{-0.2cm}
\item Bounding boxes for the printed waveforms and lead names, along with the plotted pixel coordinates of the waveforms, were recorded in metadata files alongside the generated images to train object detection models and segmentation models respectively.
\end{itemize}

@@

%----------------------------------------------------------------------------------------
%	Model selection
%----------------------------------------------------------------------------------------

\headerbox{Model Selection}{name=nn, column=2, span=1, row = 2, below=framework}{
% NOT finished

\input{tables/model_select_poster}

Models are chosen based on performance, efficiency, and compatibility with ECG image data.
\newline
\faHandPointRight[regular] DETR: 0.9 mAP \hfill \faHandPointRight[regular] U-Net: 0.73 Dice.
\newline
\faHandPointRight[regular] Total inference time: 2s on single RTX 3090

@@

%----------------------------------------------------------------------------------------
%	Novel loss functions
%----------------------------------------------------------------------------------------

\headerbox{Loss Functions}{name=loss, span=2, column=0, below=data}{
% finished

\begin{tikzpicture}
\ImageNode[]{wtmask}{width=0.4\linewidth}{images/weight-mask.pdf}
\node [rectangle, text width = 8.8cm, inner sep = 2pt, minimum height = 1.0em, right=0.3cm of wtmask.north east, anchor=north west] {\faHandPointRight[regular] Segmentation: Since ECG waveforms occupy only a small portion of the image, we used the BCE loss multiplied by a \textbf{\color{red} weight mask} to balance the loss contribution from different parts of the ECG waveforms. The weight mask is the sum of multiple dilations (e.g., dilation iterations 2, 4, 6, and 8) of the binarized ground truth mask. On the left is an example of a weight mask.};
\end{tikzpicture}
\faHandPointRight[regular] Classification: Asymmetric loss $L = -y \cdot (1-p)^{\gamma_{+}} \log(p) - (1-y) \cdot (p_m)^{\gamma_{-}} \log(1-p_m)$
\newline
\faHandPointRight[regular] Object detection: 1 $\times$ bbox class cross-entropy + 5 $\times$ bbox coordinate L1 loss + 2 $\times$ GIoU loss

@@

%----------------------------------------------------------------------------------------
%	Submission Results
%----------------------------------------------------------------------------------------

\headerbox{Submission Results}{name=submission, span=1, column=2, below=nn}{
% NOT finished

\begin{center}
\input{tables/final_results_simple}
\end{center}
The rankings are based on the hidden validation set. Results on the public training data (more specifically, the left-out validation set) are also provided for reference.

@@

%----------------------------------------------------------------------------------------
%	Limitations, Discussions
%----------------------------------------------------------------------------------------

\headerbox{Discussions and Limitations}{name=discussion, column=0, below=loss, span=3}{
% NOT finished

\begin{itemize}
\item The proposed method is relatively effective in digitizing and classifying ECG images concurrently, achieving an acceptable SNR and F1 score. This multi-stage framework offers a lightweight and efficient solution to this challenge.
\vspace{-0.2cm}
\item There are observable gaps between the SNR metric on the public training data and the hidden validation/test data. A possible reason for this discrepancy is the difference in data distribution (format): training data generated using ECG-Image-Kit all follow the standard 3-by-4 format. The post-processing step after segmentation is designed to handle this specific format. However, the hidden validation/test data may have different formats, e.g. additional full-length waveforms.
\vspace{-0.2cm}
\item Further improvement directions:
\vspace{-0.2cm}
\begin{itemize}
    \item More effective backbone networks or more advanced architectures, including the latest YOLO series models.
    \vspace{-0.1cm}
    \item End-to-end digitization models from the segmented ECG waveform masks or even directly from the ROI can be developed to simplify the pipeline, replacing the current rule-based conversion from pixel coordinates to digitized value.
    \vspace{-0.1cm}
    \item Better handling of rotated ECG images to improve the framework's robustness, using oriented object detection models.
\end{itemize}
\end{itemize}

@@

\headerbox{Acknowledgements}{name=acknowledgements, column=0, below=discussion, span=3}{
% finished
% \smaller
We thank professor {\bf Deren Han} from the School of Mathematical Sciences, Beihang University for providing GPU servers to help accomplish this work.

@@

%----------------------------------------------------------------------------------------
%	link to the GitHub repository
%----------------------------------------------------------------------------------------

\headerbox{}{name=foottext, column=0, span=3, below=acknowledgements, textborder=none, headerborder=none,  boxheaderheight=0pt}{
% finished
\hfill \small \textit{Code, configs, etc. are available at https://github.com/wenh06/cinc2025, and is based on https://github.com/DeepPSP/torch\_ecg.}
}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 2025 PhysioNet/CinC Challenge Poster
% Chagas ECG Screening - Team Revenger
% 
% Based on baposter template by T. Reveyrand
% Original template by Brian Amberg (baposter@brian-amberg.de)
% Downloaded from: http://www.LaTeXTemplates.com
% License: CC BY-NC-SA 3.0
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\usepackage{textcomp}
\usepackage{hyperref}

@@

% Color definitions (reusing 2024 poster color scheme)
\definecolor{bordercol}{RGB}{40,40,40} % Border color of content boxes
\definecolor{headercol1}{RGB}{186,215,230} % Background color for the header in the content boxes (left side)
\definecolor{headercol2}{RGB}{120,120,120} % Background color for the header in the content boxes (right side)
\definecolor{headerfontcol}{RGB}{0,0,0} % Text color for the header text in the content boxes
\definecolor{boxcolor}{RGB}{210,235,250} % Background color for the content in the content boxes

\definecolor{jdhblue}{RGB}{2,93,186}
\definecolor{zkblue}{RGB}{88,135,175}
\definecolor{zkbackground}{RGB}{230,232,234}

@@

\begin{poster}{
grid=false,
borderColor=bordercol, % Border color of content boxes
headerColorOne=headercol1, % Background color for the header in the content boxes (left side)
headerColorTwo=headercol2, % Background color for the header in the content boxes (right side)
headerFontColor=headerfontcol, % Text color for the header text in the content boxes
boxColorOne=zkbackground,
headershape=roundedright, % Specify the rounded corner in the content box headers
headerfont=\Large\sf\bf, % Font modifiers for the text in the content box headers
textborder=rectangle,
background=none,
headerborder=open, % Change to closed for a line under the content box headers
boxshade=plain
}
%
%----------------------------------------------------------------------------------------
%	TITLE AND AUTHOR NAME
%----------------------------------------------------------------------------------------
%
% TODO: Add logos when available - current paths may not exist
% \IfFileExists{images/logo_cau_science.png}{\includegraphics[scale=0.401]{images/logo_cau_science.png}}{CAU Logo}
{{\bf TODO: CAU Logo}} % University/lab logo placeholder
{
{\bf \fontsize{18pt}{18pt} \selectfont Reliability-Aware Hierarchical Learning for Chagas Detection from Electrocardiogram under Expert Label Scarcity}
} % Poster title
{\vspace{0.3em} \smaller Hao WEN$^1$, Jingsu KANG$^2$  \\  % Author names

$^1${\it College of Science, China Agricultural University}\\
$^2${\it Tianjin Medical University} \\
\vspace{0.2cm}
{\Large \bf{The George B. Moody PhysioNet Challenge 2025} $\cdot$ \bf{Team Revenger}}
}
% \IfFileExists{logo_tmu.jpeg}{\includegraphics[scale=0.101]{logo_tmu.jpeg}}{TMU Logo}
{{\bf TODO: TMU Logo}} % University logo placeholder

%----------------------------------------------------------------------------------------
%	INTRODUCTION & CHALLENGE CONTEXT
%----------------------------------------------------------------------------------------

\headerbox{Introduction \& Challenge Context}{name=introduction, column=0, row=0, span=3}{

\begin{itemize}
\item {\bf Team Revenger – PhysioNet Challenge 2025}: ECG-based Chagas cardiomyopathy screening in resource-limited settings
\item {\bf Challenge objective}: Maximize positive case retrieval under 5\% prevalence constraint for scalable screening
\item {\bf Key challenge}: Severe class imbalance ($\sim$2\% positives) + expert label scarcity (0.2\% expert-confirmed)
\item {\bf Our approach}: Reliability-aware hierarchical learning prioritizing expert-confirmed over self-reported labels
\end{itemize}

@@

%----------------------------------------------------------------------------------------
%	DATA & LABEL PROVENANCE
%----------------------------------------------------------------------------------------

\headerbox{Data \& Label Provenance}{name=data, column=0, row=1, span=2, below=introduction}{

Three ECG datasets with heterogeneous label reliability and Chagas prevalence:

\vspace{0.3cm}
\begin{center}
\begin{tabular}{rlll}
\toprule
Dataset & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Chagas \%} & Label source \\
\midrule
SaMi-Trop   & 1,631     & 100.0 \%    & expert-confirmed \\
CODE-15\%   & 345,779   & 1.8 \%      & self-reported \\
PTB-XL      & 21,799    & 0.0 \%      & N/A (non-endemic) \\
\bottomrule
\end{tabular}
\end{center}
\vspace{0.3cm}

\begin{itemize}
\item {\bf Preprocessing}: Resampled to 400 Hz, bandpass filtered (0.5–45 Hz), z-score normalized
\item {\bf Challenge}: Model must learn from heterogeneous label quality and extreme class imbalance
\end{itemize}

@@

%----------------------------------------------------------------------------------------
%	RELIABILITY-AWARE SUPERVISION
%----------------------------------------------------------------------------------------

\headerbox{Reliability-Aware Supervision}{name=supervision, column=2, row=1, span=1, below=introduction}{

{\bf Label Smoothing \& Upsampling Strategy}

Hierarchical supervision encoding source reliability through stratified label smoothing:

\vspace{0.2cm}
{\bf Formula:} $\tilde{\mathbf{y}} = (1 - \varepsilon) \cdot \mathbf{y} + \frac{\varepsilon}{C} \cdot \mathbf{1}$
\vspace{0.2cm}

Where smoothing factor $\varepsilon$ depends on reliability:
\begin{itemize}
\item SaMi positives: $\varepsilon = 0.0$ (max trust)
\item CODE-15\% samples: $\varepsilon = 0.6$ (uncertainty)  
\item PTB-XL negatives: $\varepsilon = 0.2$ (mild regularization)
\end{itemize}

\vspace{0.3cm}
\begin{center}
\begin{tabular}{ccc}
\toprule
\multicolumn{2}{c}{Upsample factor} & \multirow{2}{*}{Score} \\
\cmidrule(lr){1-2}
CODE & SaMi & \\
\midrule
3 & 12    & \textbf{0.245} \\
3 & 7     & 0.212 \\
10 & 120   & 0.210 \\
\bottomrule
\end{tabular}
\end{center}

@@

%----------------------------------------------------------------------------------------
%	MODEL ARCHITECTURE
%----------------------------------------------------------------------------------------

\headerbox{Model Architecture}{name=architecture, column=0, row=2, span=2, below=data}{

{\bf Enhanced 1D ResNet with SE Modules \& Global Pooling}

\begin{center}
\resizebox{0.95\linewidth}{!}{%
\begin{tikzpicture}[
    line width=0.75pt,
    >={Latex[length=3mm]},
    node distance=6.4mm,
    every node/.style={inner sep=2pt},
    stem/.style={draw, rounded corners=2pt, minimum width=30mm, minimum height=9.5mm,
                 align=center, fill=blue!10},
    block/.style={draw, rounded corners=2pt, minimum width=30mm, minimum height=12mm,
                  align=center, fill=orange!12},
    se/.style={draw, rounded corners=2pt, minimum width=30mm, minimum height=9.5mm,
               align=center, fill=green!12},
    pool/.style={draw, rounded corners=2pt, minimum width=30mm, minimum height=9.5mm,
                 align=center, fill=cyan!10},
    head/.style={draw, rounded corners=2pt, minimum width=30mm, minimum height=9.5mm,
                 align=center, fill=purple!10},
    output/.style={draw, rounded corners=2pt, minimum width=30mm, minimum height=9.5mm,
                align=center, fill=red!10},
    arrow/.style={-{Latex[length=3mm]}, thick},
]

% Main architecture path
\node[stem] (input) {Input \\ (12$\times$L)};
\node[stem, below=of input] (stem) {Stem \\ Conv ks=15 \\ BN + ReLU \\ 64 Ch};
\node[block, below=of stem] (bottleneck) {$4 \times$ Bottleneck Block \\ ks: 1--15--1, stride: 1--4--1 \\ Ch: $512 \to 768 \to 1024 \to 1280$};
\node[se, below=of bottleneck] (gse) {Global SE \\ reduction=8};
\node[pool, below=of gse] (gap) {Global MaxPool};
\node[head, below=of gap] (mlp) {MLP Head \\ (1280 $\rightarrow$ C=2)};
\node[output, below=of mlp] (out) {Output Logits};

% Arrows
\draw[arrow] (input) -- (stem);
\draw[arrow] (stem) -- (bottleneck);
\draw[arrow] (bottleneck) -- (gse);
\draw[arrow] (gse) -- (gap);
\draw[arrow] (gap) -- (mlp);
\draw[arrow] (mlp) -- (out);

\end{tikzpicture}
}
\end{center}

\begin{itemize}
\item {\bf Bottleneck blocks}: 1$\times$1–15$\times$1 convolutions with expansion factor 4, stride 4 downsampling
\item {\bf Global SE}: Channel attention with reduction ratio 8 for temporal feature recalibration  
\item {\bf Variable-length inputs}: Global max pooling enables inference on ECGs of any duration
\end{itemize}

@@

%----------------------------------------------------------------------------------------
%	TRAINING & OPTIMIZATION
%----------------------------------------------------------------------------------------

\headerbox{Training \& Optimization}{name=training, column=2, row=2, span=1, below=supervision}{

{\bf Asymmetric Loss \& Optimization Setup}

\begin{itemize}
\item {\bf Asymmetric Loss}: $L = -y(1-p)^{\gamma_{+}} \log(p) - (1-y)(p_m)^{\gamma_{-}} \log(1-p_m)$
\item Parameters: $(\gamma_{+},\gamma_{-})=(1,4)$, margin $\delta=0.05$
\item {\bf Optimizer}: AdamW with OneCycle scheduler
\item {\bf Learning rate}: $1 \times 10^{-4}$ initial, $6 \times 10^{-4}$ peak
\item {\bf Training}: 30 epochs, batch size 128, weight decay $1 \times 10^{-2}$
\item {\bf Input}: 4,096 samples (uniform crop/center trim)
\item {\bf Early stopping}: Challenge score on 20\% internal hold-out
\end{itemize}

@@

%----------------------------------------------------------------------------------------
%	RESULTS
%----------------------------------------------------------------------------------------

\headerbox{Results}{name=results, column=0, row=3, span=3, below=architecture}{

{\bf Challenge Performance \& Analysis}

\begin{center}
\begin{tabular}{r|r|r|r}
Internal hold-out (mean±std) & Hidden validation & Rank & Test \\
\hline
$0.451 \pm 0.005$ & 0.245 & 179/364 & TBA \\
\hline
\end{tabular}
\end{center}

\begin{itemize}
\item {\bf Hidden validation score}: 0.245 (Challenge metric: TPR at 5\% prevalence constraint)
\item {\bf Performance gap}: Internal vs. hidden suggests domain shift or overfitting to public data
\item {\bf Key insight}: Reliability-aware supervision stabilizes learning under noisy labels more effectively than architectural complexity alone
\end{itemize}

%----------------------------------------------------------------------------------------
%	DISCUSSION & INSIGHTS
%----------------------------------------------------------------------------------------

\headerbox{Discussion \& Insights}{name=discussion, column=0, row=4, span=2, below=results}{

{\bf How does the approach provide insight into the Challenge issues?}

\begin{itemize}
\item {\bf Label provenance modeling}: Explicitly encoding expert vs. self-reported reliability addresses core challenge of heterogeneous label quality in multi-source ECG datasets
\item {\bf Hierarchical supervision}: Dynamic smoothing factors prevent overconfident gradients on noisy labels while preserving sharp supervision on expert-confirmed cases
\item {\bf Scalability}: Framework aligns with clinical reality—expert labels are scarce but high-quality; self-reported labels are abundant but uncertain
\item {\bf Performance gap analysis}: Static reliability weights may not capture within-source label heterogeneity; adaptive weighting could improve generalization
\end{itemize}

%----------------------------------------------------------------------------------------
%	FUTURE WORK
%----------------------------------------------------------------------------------------

\headerbox{Future Work}{name=future, column=2, row=4, span=1, below=results}{

{\bf Adaptive Supervision Framework}

\begin{itemize}
\item {\bf Dynamic reliability}: Learn instance-specific confidence scores vs. static smoothing factors
\item {\bf Adaptive sampling}: Confidence-aware upsampling responding to model evolution during training
\item {\bf Data augmentation}: CutMix, SMOTE for positive sample diversification
\item {\bf Multi-task learning}: Leverage auxiliary arrhythmia labels for enhanced feature representation
\item {\bf Self-supervised pre-training}: Large-scale unlabeled ECG data for transferable representations
\end{itemize}

@@

\headerbox{Acknowledgements}{name=acknowledgements, column=0, row=5, span=3, below=discussion}{

We thank Professor {\bf Deren Han} from the School of Mathematical Sciences, Beihang University for providing GPU servers. We acknowledge the PhysioNet Challenge organizers and the contributing datasets.

@@

%----------------------------------------------------------------------------------------
%	REFERENCES
%----------------------------------------------------------------------------------------

\headerbox{References}{name=references, column=0, row=6, span=3, below=acknowledgements}{

\begin{thebibliography}{9}
\bibitem{goldberger2000} Goldberger AL, et al. PhysioBank, PhysioToolkit, and PhysioNet. {\it Circulation} 2000;101(23):e215-e220.
\bibitem{ribeiro2020} Ribeiro AH, et al. Automatic diagnosis of the 12-lead ECG using a deep neural network. {\it Nat Commun} 2020;11:1760.
\bibitem{hu2018} Hu J, et al. Squeeze-and-Excitation Networks. {\it CVPR} 2018:7132-7141.
\bibitem{ridnik2021} Ridnik T, et al. Asymmetric Loss For Multi-Label Classification. {\it ICCV} 2021:82-91.
\bibitem{yun2019} Yun S, et al. CutMix: Regularization Strategy to Train Strong Classifiers. {\it ICCV} 2019:6023-6032.
\bibitem{chawla2002} Chawla NV, et al. SMOTE: Synthetic Minority Oversampling Technique. {\it JAIR} 2002;16:321-357.
\end{thebibliography}

%----------------------------------------------------------------------------------------
%	FOOTER
%----------------------------------------------------------------------------------------

\headerbox{}{name=foottext, column=0, span=3, below=references, textborder=none, headerborder=none, boxheaderheight=0pt}{
\hfill \small \textit{Code, configs, etc. are available at https://github.com/wenh06/cinc2025, and is based on https://github.com/DeepPSP/torch\_ecg.}
}
